---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
We begin in a working directory that contains the activity file and read in the contents of the file. 

```{r}
setwd("~/Desktop/R Data/RepData_PeerAssessment1")

dat <- read.csv("activity.csv")
```

Take a peek at the contents of dat.
```{r}
head(dat)
```
```{r}
tail(dat)
```
```{r}
str(dat)
```

The activity CSV contains a number of steps during a given five-minute period. For each date, the interval variable is the beginning of the five-minute portion of the day (in military time). 
## What is mean total number of steps taken per day?
To begin, we have to find the mean sum of day's steps.
```{r}
library(dplyr)
library(ggplot2)
dat2 <- dat %>% group_by(date) %>% summarize(Daily_Steps=sum(steps))
dat2

g <- ggplot(data=dat2, aes(x= date, y= Daily_Steps))+
        geom_bar(stat="identity")+
        theme(axis.title.x = element_text(face="bold", colour="#990000", size=20),
           axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
print(g)

```

From here, it's simple to find the average of all the days.

```{r}
mean(dat2$Daily_Steps, na.rm=TRUE) #na.rm=TRUE is important or else this will return NA
```

## What is the average daily activity pattern?
Finding the daily activity pattern means we have to start back at dat and summarize the steps by interval.
```{r}

```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
